Effects of Different Pain Relievers and Dosages on Cognitive Retention

### 2025-04-08 paper/paper.qmd
Update paper.qmd
Effect Size and Power Analysis

We did a post-hoc power analysis. Based on the overall $\eta^2$ from the ANOVA model, we calculated a Cohen's $f$ value of 0.20. This suggests a small-medium effect size, which means that the type of drug and dosage level may still have had a meaningful impact on memory task performance.

The estimated power of the study was around 59.6%. This means that our study was moderately able to detect an effect with the size we have, if there was an effect. Despite being less than the ideal 80% power value, this suggests that the study is still able to detect large & meaningful differences -- just not smaller ones.

@@

```{r}
# EFFECT SIZE AND POWER ANALYSIS
eta_sq <- eta_squared(anova_model, partial = FALSE)

etasquare <- sum(eta_sq$Eta2)  # sum of all eta^2 values
cohens_f <- sqrt(etasquare / (1 - etasquare))

power_result <- wp.kanova(
  n = 240,
  ndf = 7, # because 8 - 1 = 7
  f = cohens_f,
  ng = 8, # because 8 treatment groups
  alpha = 0.05
)

>>>>

We did a post-hoc power analysis. Based on our two-way ANOVA model, we calculated the power for the two main effects and the interaction effect (ie. drug, dosage_level, and drug:dosage_level, respectively).

We achieved the following effect sizes (Cohen's f) and powers. For drug type, we got the effect size of 0.134 and a power of 37.6%. For the dosage level, we got the effect size of 0.074 and a power of 20.8%. And for the interaction effect for the drug and dosage level, we got the effect size of 0.135 and a power of 38.2%.

These results show that the statistical power for all three effects were low. In other words, our study had a limited ability to detect real effects, should they exist. With power values of 37.6%, 20.8%, and 38.2% for the drug, dosage level, and interaction effects respectively, we had less than a 40% chance of identifying a true effect under the conditions of our study.

This suggests that while our ANOVA results were not statistically significant, this may be due to insufficient power rather than the absence of any real effect. The observed effect sizes fall into the small range according to Cohen’s conventions, which typically require a larger sample size to detect reliably.

We discuss about addressing this limitation in the next section.

@@

```{r}
# EFFECT SIZE AND POWER ANALYSIS
eta_sq <- eta_squared(anova_model, partial = TRUE)

eta2drug <- eta_sq$Eta2[1]  
cohenfdrug <- sqrt(eta2drug / (1 - eta2drug))
print(cohenfdrug)

eta2dos <- eta_sq$Eta2[2]  
cohenfdos <- sqrt(eta2dos / (1 - eta2dos))
print(cohenfdos)

eta2drugdos <- eta_sq$Eta2[3]  
cohenfdrugdos <- sqrt(eta2drugdos / (1 - eta2drugdos))
print(cohenfdrugdos)

powerdrug <- wp.kanova(
  n = 240,
  ndf = 3,
  f = cohenfdrug,
  ng = 8,
  alpha = 0.05
)

powerdos <- wp.kanova(
  n = 240,
  ndf = 1, 
  f = cohenfdos,
  ng = 8,
  alpha = 0.05
)

powerdrugdos <- wp.kanova(
  n = 240,
  ndf = 3,
  f = cohenfdrugdos,
  ng = 8,
  alpha = 0.05
)

print(powerdrug)
print(powerdos)
print(powerdrugdos)
```
